---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
file.remove(list.files("R", pattern = "day.*", full.names = TRUE))
file.remove(list.files("inst", pattern = "input.*", full.names = TRUE))
file.remove(list.files("inst", pattern = "run.*", full.names = TRUE))
file.remove(list.files("tests/testthat", pattern = "test.day*", full.names = TRUE))
file.remove("R/data-solutions.R")
file.rename(".aoccookie", ".xaoccookie")
aoc::use_day(01, 2020, open = FALSE)
```

# aoc

<!-- badges: start -->
<!-- badges: end -->

aoc provides [usethis](https://usethis.r-lib.org/)-style functions for [Advent
of Code](https://adventofcode.com) puzzles. This package only downloads content
from the Advent of Code site.

## Installation

You can install the development version of aoc from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
remotes::install_github("tjmahr/aoc")
```

## Preliminaries

aoc assumes that we are organizing our R code using an R package. Therefore, it
requires a minimal package infrastructure in order to work. In RStudio, using

  - File \> New Project... \> New Directory \> R Package
  - Setting the package name to `adventofcode21` \> Create Project

should create enough of an R package setup for aoc to work.

## Functionality

aoc assumes that you are working inside an R package. By default, it assumes the
package is named `adventofcodeXX` where XX are the last two digits of the year.
For example, `adventofcode20` would be the package for 2020.

From this setup, aoc will automate a number of tasks. `use_day()` is a
[usethis](https://usethis.r-lib.org/)-style function to create placeholder files
for each day.

`use_day(day = 1)` does the following tasks:

  - on first run, creates `R/data-solutions.R`
    - a file for storing solutions
    
  - creates `R/day01.R`
    - puts the puzzle description into a roxygen documentation block
    - creates placeholder functions for the solutions to day 1
    
  - creates `tests/testthat/test-day01.R`
    - a unit test for day (useful for the example in the puzzle
      description) in `tests/testthat/test-day01.R`
      
  - creates `inst/input01.txt`
    - a file to hold the input for day 1
    
  - creates `inst/run-day01.R` 
    - a file to contain the solution for day 1

```{r, eval = FALSE}
aoc::use_day(1, year = 2020)
#> √ Writing 'R/data-solutions.R'
#> * Modify 'R/data-solutions.R'
#> Executing: pandoc -t markdown -o 
#> "C:\Users\trist\AppData\Local\Temp\RtmpiYf4Vi\file42a83cf05e25.markdown" 
#> ">C:\Users\trist\AppData\Local\Temp\RtmpiYf4Vi\file42a83cf05e25.html"
#> √ Writing 'R/day01.R'
#> * Modify 'R/day01.R'
#> ● Write your solution code here
#> ● Once you unlock Part Two, update the Roxygen block with the description
#> √ Writing 'inst/input01.txt'
#> * Modify 'inst/input01.txt'
#> ● Copy your problem input into this file
#> √ Writing 'tests/testthat/test-day01.R'
#> * Modify 'tests/testthat/test-day01.R'
#> ● Write unit tests using the examples from the problem description
#> √ Writing 'inst/run-day01.R'
#> * Modify 'inst/run-day01.R'
#> ● Run your solution on the input here. Once it works, update R/data-solutions.R
```

`R/data-solutions.R` is where we store our solutions:

```{r, echo = FALSE}
preview_lines <- function(path, lines = 10) {
  # path <- "R/utils-usethis.R"
  l <- readLines(path)
  if (lines * 2 < length(l)) {
    show <- c(head(l, lines), "[... truncated ...]")
  } else {
    show <- l
  }
  writeLines(show)
}
```


```{r, echo = FALSE, comment = ""}
preview_lines("R/data-solutions.R", 20)
```

`inst/input01.txt` is an empty file for our input data. We have to paste in our
puzzle input here.

The R script for `R/day01.R` provides the puzzle description for part 1,
function stubs for part 1 `f01a()` and part 2 `f01b()`. I also like to make the
example data into a function for unit tests or code examples, so there is a stub
for `example_data_01()`.

```{r, echo = FALSE, comment = ""}
preview_lines("R/day01.R", 200)
```

`tests/testthat/test-day01.R` is a placeholder for file unit tests. It's a good
place work through the examples in the puzzle description.

```{r, echo = FALSE, comment = ""}
preview_lines("tests/testthat/test-day01.R")
```

Once we have developed a solution for the example input, we can test our
official input by running the code in `inst/run-day01.R`. The final two lines
provide code to validate the solutions that we store in `R/data-solutions.R`.

```{r, echo = FALSE, comment = ""}
preview_lines("inst/run-day01.R")
```



### Default values for `use_day()`

We can tell aoc which year to use by using `options()`.

```{r, eval = TRUE}
options(aoc.year = 2017)
aoc::use_day(3)
```

If we look at the first lines of `R/day03.R`, we can see the correct URL used.

```{r, echo = FALSE, comment = ""}
preview_lines("R/day03.R")
```

We can also tell aoc which package name to use for our project using
`options()`.

```{r, eval = TRUE}
options(aoc.package = "awesomeadvent2017")
aoc::use_day(4)
```

And here the correct name appears in the `library()` call.

```{r, echo = FALSE, comment = ""}
preview_lines("inst/run-day04.R")
```

We can set these permanently for an Advent of Code package by editing the
package's `.Rprofile`:

```{r, eval = FALSE}
usethis::edit_r_profile(scope = "project")
# add in things like `options(aoc.year = 2017)`
```

### Advanced: Using a user cookie 

If you know how to retrieve the cookie for your Advent of Code user, you can use
this cookie to download your puzzle input. Store the cookie in file named
`.aoccookie`. Then `use_day()` will automatically use this cookie when
downloading puzzle input.

For these demos, I hid my cookie by renaming the file. If I unrename the file
and download the day 7 files, I can preview the lines of the input file.

```{r}
file.rename(".xaoccookie", ".aoccookie")

aoc::use_day(7)

# this is a function i defined in a hidden code block 🤫
preview_lines("inst/input07.txt")
```

The other advantage of a user cookie is that after solving part 1 of a day, we
can download part 2 as a roxygen2 block. By default, this block is copied to the 
clipboard, but for this demo, I have to disable it.

```{r, comment = ""}
aoc::download_part2_to_roxygen_md(day = 7, clip = FALSE)
```


```{r, include = FALSE}
file.remove(list.files("R", pattern = "day.*", full.names = TRUE))
file.remove(list.files("inst", pattern = "input.*", full.names = TRUE))
file.remove(list.files("inst", pattern = "run.*", full.names = TRUE))
file.remove(list.files("tests/testthat", pattern = "test.day*", full.names = TRUE))
file.remove("R/data-solutions.R")
```

